Simultaneous Elicitation of Scoring Rule and Agents Preferences for Robust Winner Determination

### 2020-01-21 draft.tex
Renamed strats

Let $V$ be the set of possible preference profiles (the union, for any integer, $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
%\commentBN{$s^{\profile,\w}(x)$}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with highest scores.

@@

%TODO Limited; chair
The \strat{LimitedPessimistic} strategy uses the same criterion as the pessimistic strategy, but only considering a small set of candidate questions.
For each voter, a candidate question is chosen by considering the heuristic proposed by \citet{Lu2011}. 
For the chair, we define
\begin{align}
w^\tau & = \argmin_{\w \in W} s(\bar{y}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w), \label{eq:tau} 
\end{align}
and compare component-wise $\bar{\w}$ and $\w^{\tau}$, to find the position $r$ that is most valuable to ask about (the one that maximizes $\abs{\bar{\w}_r - \w^{\tau}_r}$), and choose $\lambda$ as the middle of the interval of possible values for that rank.
\commentOC{I am not so happy about this. If we have time, we should consider one question per rank and drop the tau business.}
We then evaluate, for each candidate question, the minimax regret for each possible answer; finally we choose the best according to $\leximax$.
%\commentOC{Reducing the complexity to $O(n^2 m^3). Is this right? If we can solve easily a size $(m, n) = (15, 15)$ with the Pwh strategy, which requires about $15^5 \approx 8 . 10^5$ operations, we should be able to solve easily a size $(m, n) = (6, 6)$ with the Pessimistic strategy, which requires about $6^7 \approx 3 . 10^5$ operations.}

Finally, the \strat{Two phase} strategy is developed in order to investigate the effect of varying the proportion of questions asked to voters and to the committee. The \strat{Two phase - vc} first asks $p$ questions to the voters, then $k-p$ questions to the committee, using in both cases Pessimistic with heuristic \commentBN{limitedPessimistic?} to select the specific question. The \strat{Two phase - cv} strategy starts with $p$ questions to the committee, then asks to the voters. \commentOC{If our XPs exhibit no significant change between vc and cv (or if we don’t have time to run them), I will drop this distinction.}

@@

\sisetup{round-mode=places, round-precision=1, table-figures-integer=1, table-figures-decimal=1}
\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {Pes.} & {sd} & {Pes.\ w.\ h.} & {sd} \\
			\midrule
			0 & 5.0 & 0 & 5.0 & 0\\
			5 & 4.93 & 0.16 & 4.15 & 0.32\\
			10 & 4.84 & 0.22 & 3.44 & 0.46\\
			15 & 4.28 & 0.58 & 2.66 &0.54 \\
			20 & 3.85 & 0.33 & 1.62 & 0.82\\
			25 & 3.50 & 0.78 & 0.84 &  0.57\\
			30 & 2.96 & 0.80 & 0.53 & 0.70\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Minimax regret in problems of size $(5, 5)$ after $k$ questions.}
	\label{fig:55}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {Pes.\ w.\ h.} & {sd} \\
			\midrule
			0 & 6.0 & 0 &  &  \\
			5 &  &  &  &  \\
			10 &  &  &  &  \\
			15 &  &  &  &  \\
			20 &  &  &  &  \\
			25 &  &  &  &  \\
			30 &  &  &  &  \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Regret in problems of size $(10, 20)$ after $k$ questions.}
	\label{fig:1020}
\end{table}

We test our strategies using randomly generated datasets. %TODO and real datasets. (sushi and iris).
Our first goal is to see, with a small problem size ($m = 5, n = 5$) and the (time consuming) Pessimistic strategy, if an important lowering of the maximal regret can be achieved with a reasonable number of questions. We also want to estimate how “hard” such a problem is, by using the Random strategy as a baseline. Third, we want to estimate the loss (in terms of worst regret) when switching to the faster Pessimistic with heuristic strategy. Fourth, we want to see, on a bigger problem size ($m = 10, n = 20$), how many questions must be asked for that strategy to achieve a significant reduction of the worst regret. Finally, we want to evaluate, thanks to the Two phases strategy, the impact of varying the proportion of questions asked to the voters (with respect to questions asked to the committee) on the reduction of the worst regret.
%We also want to see how realistic our approach is for various problem sizes, by estimating the number of questions we would have to ask before being able to reduce the worst regret significantly.

@@

The results are displayed in \cref{fig:55,fig:1020,fig:p}, where strategies are designated by Rnd for Random, 2 ph. for Two phases, Pes. for Pessimistic and Pes.\ w.\ h. for Pessimistic with heuristic.

>>>>

Let $V$ be the set of possible preference profiles (the union, for any integer, $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
%\commentBN{$s^{\profile,\w}(x)$}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives gighest scores.

@@

%TODO Limited; chair
The \strat{Limited pessimistic} strategy uses the same criterion as the pessimistic strategy, but only considering a small set of candidate questions.
For each voter, a candidate question is chosen by considering the heuristic proposed by \citet{Lu2011}. 
For the chair, we define
\begin{align}
w^\tau & = \argmin_{\w \in W} s(\bar{y}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w), \label{eq:tau} 
\end{align}
and compare component-wise $\bar{\w}$ and $\w^{\tau}$, to find the position $r$ that is most valuable to ask about (the one that maximizes $\abs{\bar{\w}_r - \w^{\tau}_r}$), and choose $\lambda$ as the middle of the interval of possible values for that rank.
\commentOC{I am not so happy about this. If we have time, we should consider one question per rank and drop the tau business.}
We then evaluate, for each candidate question, the minimax regret for each possible answer; finally we choose the best according to $\leximax$.
%\commentOC{Reducing the complexity to $O(n^2 m^3). Is this right? If we can solve easily a size $(m, n) = (15, 15)$ with the Pwh strategy, which requires about $15^5 \approx 8 . 10^5$ operations, we should be able to solve easily a size $(m, n) = (6, 6)$ with the Pessimistic strategy, which requires about $6^7 \approx 3 . 10^5$ operations.}

Finally, the \strat{Two phases} strategy is developed in order to investigate the effect of varying the proportion of questions asked to voters and to the committee. The \strat{Two phases - vc} first asks $p$ questions to the voters, then $k-p$ questions to the committee, using in both cases Limited pessimistic to select the specific question. The \strat{Two phases - cv} strategy starts with $p$ questions to the committee, then asks to the voters. \commentOC{If our XPs exhibit no significant change between vc and cv (or if we don’t have time to run them), I will drop this distinction.}

@@

\sisetup{round-mode=places, round-precision=1, table-figures-integer=1, table-figures-decimal=1}
\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {Pes.} & {sd} & {L.\ pes.} & {sd} \\
			\midrule
			0 & 5.0 & 0 & 5.0 & 0\\
			5 & 4.93 & 0.16 & 4.15 & 0.32\\
			10 & 4.84 & 0.22 & 3.44 & 0.46\\
			15 & 4.28 & 0.58 & 2.66 &0.54 \\
			20 & 3.85 & 0.33 & 1.62 & 0.82\\
			25 & 3.50 & 0.78 & 0.84 &  0.57\\
			30 & 2.96 & 0.80 & 0.53 & 0.70\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Minimax regret in problems of size $(5, 5)$ after $k$ questions.}
	\label{fig:55}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {L.\ pes.} & {sd} \\
			\midrule
			0 & 6.0 & 0 &  &  \\
			5 &  &  &  &  \\
			10 &  &  &  &  \\
			15 &  &  &  &  \\
			20 &  &  &  &  \\
			25 &  &  &  &  \\
			30 &  &  &  &  \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Regret in problems of size $(10, 20)$ after $k$ questions.}
	\label{fig:1020}
\end{table}

We test our strategies using randomly generated datasets. %TODO and real datasets. (sushi and iris).
Our first goal is to see, with a small problem size ($m = 5, n = 5$) and the (time consuming) Pessimistic strategy, if an important lowering of the maximal regret can be achieved with a reasonable number of questions. We also want to estimate how “hard” such a problem is, by using the Random strategy as a baseline. Third, we want to estimate the loss (in terms of worst regret) when switching to the faster Limited pessimistic strategy. Fourth, we want to see, on a bigger problem size ($m = 10, n = 20$), how many questions must be asked for that strategy to achieve a significant reduction of the worst regret. Finally, we want to evaluate, thanks to the Two phases strategy, the impact of varying the proportion of questions asked to the voters (with respect to questions asked to the committee) on the reduction of the worst regret.
%We also want to see how realistic our approach is for various problem sizes, by estimating the number of questions we would have to ask before being able to reduce the worst regret significantly.

@@

The results are displayed in \cref{fig:55,fig:1020,fig:p}, where strategies are designated by Rnd for Random, 2 ph. for Two phases, Pes. for Pessimistic and L.\ pes. for Limited pessimistic.
